---
title: 'Capstone Week 3: Modeling'
author: "JRB"
date: "November 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(tidy = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(fig.align = "center")
  
library(tidyverse) ##Because the universe is a better place with tidyverse
library(tidytext)
library(tm)
library(ggraph)
library(igraph)
library(knitr) #For table formatting
library(gridExtra)
library(ggthemes)
## Formating for plots
ggstyle <- function() {
  theme_tufte() +
  theme(
                text = element_text(
                                        face="italic",
                                        colour="black", 
                                        size=10),
                plot.title = element_text(
                              face="bold",
                              colour="black",
                              size=11,
                              hjust = 0.5
                )
                        )

} 
## Formatting for graphs
a <- grid::arrow(angle=45,
                 type="open",
                 unit(4,"mm"))
```
```{r}
twitter<- readLines("./final/en_US/en_US.twitter.txt")
twitter_df <- data_frame(corp="twitter",line=1:length(twitter),text=twitter)
corp <- twitter_df
## This is is our "profanity" list
swears <- readLines("./data/profanity.txt")
## Prepping our list of swear words
swears <- as_tibble(as.character(swears))
colnames(swears) <- "word"

```
```{r}
#Step 1: subset for twitter
#corp <-twitter_df
#corp <- sample_n(corp,20000)
#Step 2: normalizing case, cleaning twitter artifacts and removing digits
corp$text <- tolower(corp$text)
corp$text <- gsub("rt","",corp$text) #Remove rt, retweet
corp$text <- gsub("@\\w+", "", corp$text) #Remoeve any @
corp$text <- gsub("[[:digit:]]", "", corp$text) # remove numbers

#Step 3: removing stop words (a, the, etc...) and profanity
data(stop_words)
tidy_corp_nostop <- corp %>%
                unnest_tokens(word,text,token="words") %>%
                anti_join(swears, by="word") %>%
                anti_join(stop_words, by="word")

tidy_corp_stop <- corp %>%
                unnest_tokens(word,text,token="words") %>%
                anti_join(swears, by="word")

tidy_corp_nostop_summary <- tidy_corp_nostop %>% 
  group_by(corp) %>%
  count(word, sort = TRUE) %>%
  mutate(word = reorder(word, n)) %>%
  bind_tf_idf(word,corp,n) 

#tidy_corp_nostop_summary <- tidy_corp_nostop_summary %>%  mutate(cumsum=cumsum(n)) %>% mutate(cumsumfreq=cumsum(tf))

tidy_corp_stop_summary <- tidy_corp_stop %>%
  group_by(corp) %>%
  count(word, sort = TRUE) %>%
  mutate(word = reorder(word, n)) %>%
  bind_tf_idf(word,corp,n)
```
```{r 3-grams}
corp_trigram <- unnest_tokens(corp[corp$corp=="twitter",],ngram,text,token="ngrams",n=3)
corp_trigram <- corp_trigram %>%
          separate(ngram,c("word1","word2","word3"),sep=" ")
corp_trigram_count <- corp_trigram %>%
          unite(trigram,word1, word2,word3, sep=" ") %>%
          count(trigram, sort=TRUE) %>%
          mutate(orig="twitter") %>%
          arrange(desc(n))
corp_trigram_count <- corp_trigram_count %>% 
                    bind_tf_idf(trigram,orig,n) %>%
                    mutate(trigram=factor(trigram,levels=trigram))

g1 <-  corp_trigram_count %>% 
  filter(rank(desc(tf))<= 25) %>%
  ggplot(aes(reorder(trigram,desc(trigram)),tf)) +
  geom_point(stat="identity") + 
  xlab(NULL) +
  coord_flip()+
  ylab("Frequency") +
  ggtitle("25 most frequent trigrams") +
    theme_minimal() +
  theme(
                text = element_text(
                                        face="italic",
                                        colour="black", 
                                        size=8)
)
g1

trigram_graph <- corp_trigram %>% 
                count(word1,word2,word3,sort=TRUE) %>%
                filter(rank(desc(n)) <= 125) %>%
                graph_from_data_frame() 
#a <- grid::arrow(type="closed",unit(0.15,"inches"))
ggraph(trigram_graph, layout = "fr") +
      geom_edge_link(aes(edge_alpha=n),
                  show.legend = FALSE,
                  arrow=a,end_cap=circle(0.07,"inches") ) +
      geom_node_point(color="lightblue",size=5) +
      geom_node_text(aes(label = name), vjust = 1, hjust = 1) + 
      theme_graph() +
      ggtitle("Connected graph showing the 125 most frequents trigrams") +
      theme(
        plot.title= element_text(
          face="bold",
          size=10,
          hjust=0.5
        )
      )


